---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Download and save the zipfile in the working directory and rename zipfile "activity"
```{r, echo=TRUE}
unzip(zipfile="activity.zip")
activity<-read.csv("activity.csv", header=TRUE)
activity$date<-as.Date(activity$date, format="%Y-%m-%d")
```

### Exploratory Analysis
```{r, echo=TRUE}
head(activity)
```

```{r, echo=TRUE}
tail(activity)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
stepsperday<-tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

### Histogram of the total number of steps taken per day
```{r, echo=TRUE}
hist(stepsperday, breaks=20)
```

### Mean of the total steps taken per day
```{r, echo=TRUE}
meanstepsperday<-mean(stepsperday)
meanstepsperday
```

### Median of the total steps taken per day
```{r, echo=TRUE}
medianstepsperday<-median(stepsperday)
medianstepsperday
```

## What is the average daily activity pattern?
### Calculating the average number of steps taken per day
```{r, echo=TRUE}
avgstepsperday<-aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```

### Time series plot of average number of steps taken
```{r, echo=TRUE}
library(ggplot2)
g<-ggplot(data=avgstepsperday, aes(x=interval, y=steps))+geom_line()
g<-g+xlab("5 minute intervals")+ ylab("average number of steps")
g
```

### Calculating the five minute interval that contains the maximum number of steps taken
```{r, echo=TRUE}
avgstepsperday[which.max(avgstepsperday$steps),]
```

## Imputing missing values
### Counting the number of missing values
```{r, echo=TRUE}
nrow(activity[is.na(activity$steps),])
```

### Imputing missing data
Missing values are imputed using the mean through the impute function in the Hmisc package.
```{r, echo=TRUE}
library(Hmisc)
imputedactivity<-activity
imputedactivity$steps<-impute(activity$steps, fun=mean)
```

### Calculating the total steps per day using the imputed data
```{r, echo=TRUE}
imputedstepsperday<-tapply(imputedactivity$steps, imputedactivity$date, sum)
```

### Histogram of total steps per day using imputed data
```{r, echo=TRUE}
hist(imputedstepsperday, breaks=20)
```

### Mean of total steps per day using imputed data
```{r, echo=TRUE}
mean_imputed_steps_per_day<-mean(imputedstepsperday)
mean_imputed_steps_per_day
```

### Median of total steps per day using imputed data
```{r, echo=TRUE}
median_imputed_steps_per_day<-median(imputedstepsperday)
median_imputed_steps_per_day
```
We can see that there is a difference between the measures before and after imputing the data. Imputing missing data gives higher values for both mean and median.

## Are there differences in activity patterns between weekdays and weekends?

### Constructing a factor variable with two levels: "weekend" and "weekday" that indicates whether a given day is a weekday or a weekend.
```{r, echo=TRUE}
imputedactivity$day<-ifelse(as.POSIXlt(imputedactivity$date)$wday %in% c(0,6), 'weekend','weekday')
```

### Calculating the average steps per day from the imputed data
```{r, echo=TRUE}
avg_steps_imputedactivity<-aggregate(steps~interval+day, data=imputedactivity, mean)
```

### Panel plot of average steps comparing weekdays and weekends
```{r, echo=TRUE}
g1<-ggplot(avg_steps_imputedactivity,aes(interval, steps))+geom_line()+facet_grid(day~.)+xlab("5 minute interval")+ylab("mean steps taken")
g1
```


